```{r setupvarpartth, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
theme_set(bayesplot::theme_default())
knitr::opts_chunk$set(
	echo = F,
	fig.height = 6,
	fig.width = 8,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	cache.lazy = F
)
```

```{r varpartdata1}
S <- 100
I <- 100
sigma_sp <- 1
sigma_ind <- 1
data <- data.frame(species = 1:S, trait = rnorm(S, 0, sigma_sp)) %>% 
  group_by(species) %>% 
  do(data.frame(individual = 1:I, trait = rnorm(I, .$trait, sigma_ind))) %>% 
  mutate_at(c("species", "individual"), as.factor)
subdata1 <- data %>% 
  filter(species %in% unique(data$species)[1:25]) %>% 
  group_by(species) %>% 
  sample_n(4) %>% 
  ungroup()
subdata2 <- data %>% 
  filter(species %in% unique(data$species)[1:4]) %>% 
  group_by(species) %>% 
  sample_n(25) %>% 
  ungroup()
subdata3 <- data %>% 
  filter(species %in% unique(data$species)[1:10]) %>% 
  group_by(species) %>% 
  sample_n(10) %>% 
  ungroup()
data <- list("full" = data, 
             "unbalanced species" = subdata1, 
             "unbalanced individuals" = subdata2, 
             "balanced" = subdata3) %>% 
  bind_rows(.id = "sampling") %>% 
  mutate(sampling = factor(sampling, levels = c("full", "unbalanced species",
                                                "unbalanced individuals", "balanced")))
```

# Balance sampling 

This chapter demonstrates the importance of a balanced design in the distribution of variance, in contrast to the many papers reporting inter- and intraspecific variation in unbalanced designs. It also virtually explores the evaluation of leaf and measurement variation.

## Species and individual sampling

We simulated 100 species with a species variance of 1 for a given trait with 100 individuals per species with a variance of 1 for the same trait. We tested the effect of four sampling strategies on variance partitioning (Fig. \@ref(fig:traitdist)):

1. full sampling of the community
1. sampling of 100 individuals unbalanced in species (25 species with 4 individuals)
1. sampling of 100 individuals unbalanced in individuals (4 species with 25 individuals)
1. sampling of 100 individuals unbalanced in species and individual (10 species with 10 individuals)

The results demonstrate the need to use a balanced sampling design to properly assess and compare functional variation between and within species (Fig. \@ref(fig:varpart)).

```{r traitdist, fig.cap="Trait distribution per species with balanced and unbalanced sampling design."}
ggplot(data, aes(trait, group = species)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ sampling, scales = "free")
```

```{r varpart, fig.cap="Effect of the sampling strategy on variance partitioning. Variance partitioning was obtained using linear mixed models for every sampling strategy: 4 individuals in 25 species (unbalanced species), 25 individuals in 4 species (unbalanced individuals), and 10 individuals in 10 species (balanced). Dashed line represents the expected variance partitioning based on the full community of 100 individuals in 100 species."}
varcols <- c(species = "red", individual = "lightgreen", 
             sample = "lightblue", repetition = "orange", residual = "grey")
fun <- function(data){
  var <- ape::varcomp(nlme::lme(trait ~ 1, random=~1|species/individual, data), 1)
  return(data.frame(level = names(var), variance = as.vector(var)))
}
varpart <-  data %>% 
  group_by(sampling) %>% 
  do(var = fun(.)) %>% 
  unnest(var) %>% 
  mutate(level = recode(level, "Within" = "residual")) %>% 
  mutate(level = factor(level, levels = c("species", "individual", "residual")))

  
g <- ggplot(filter(varpart, sampling != "full"),
       aes(x = sampling, y = variance, fill = level)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_hline(aes(yintercept = variance), linetype = "dashed",
             data = filter(varpart, sampling == "full", level != "species") %>% 
               summarise(variance = sum(variance))) +
  scale_fill_manual(expression(sigma^2),
                    values = unname(varcols[c("species", "individual", "residual")])) +
  ylab("Percentage of variance") + xlab("Sampling strategy")
# ggsave(filename = "~/Téléchargements/Fig1.png", plot = g, bg = "white", height = 4, width = 6)
g
```

## Leaf variations and measurement errors

Similarly, we examined a sampling of 10 individuals in 10 species, 10 leaves for each individual, and 10 measurements for each leaf. This would allow us to assess the level of variation in the measurement alone to avoid an unrealistic sampling of 10 species times 10 individuals times 10 leaves times 10 measurements, which would yield 10,000 measurements! We found very stable measurements across levels using very different variations between levels (Fig. \@ref(fig:varpart2)).

```{r varpartdata2}
S <- 10
I <- 10
L <- 10
M <- 10
sigma_sp <- 6
sigma_ind <- 4
sigma_leaf <- 2
sigma_meas <- 1
data <- data.frame(species = 1:S, trait = rnorm(S, 0, sigma_sp)) %>% 
  group_by(species) %>% 
  do(data.frame(individual = 1:I, trait = rnorm(I, .$trait, sigma_ind))) %>% 
  group_by(species, individual) %>% 
  do(data.frame(leaf = 1:L, trait = rnorm(L, .$trait, sigma_leaf))) %>% 
  group_by(species, individual, leaf) %>% 
  do(data.frame(measure = 1:M, trait = rnorm(M, .$trait, sigma_meas))) %>% 
  ungroup() %>% 
  mutate_at(c("species", "individual", "leaf", "measure"), as.factor)
subdata <- bind_rows(
  # 10 sps & 10 inds
  data %>% 
    filter(leaf == 1) %>% 
    filter(measure == 1),
  # 10 inds & 10 leaves
  data %>% 
    filter(species == 1) %>% 
    filter(measure == 1),
  # 10 leaves & 10 measures
  data %>% 
    filter(species == 1) %>% 
    filter(individual == 1)
)
```

```{r varpart2,  fig.cap="Variance partitioning with balanced sampling design across levels."}
fun <- function(data){
  var1 <- ape::varcomp(nlme::lme(trait ~ 1, random=~1|species/individual, data), 1)
  var2 <- ape::varcomp(nlme::lme(trait ~ 1, random=~1|individual/leaf, 
                                 filter(data, species == 1)), 1)
  var3 <- ape::varcomp(nlme::lme(trait ~ 1, random=~1|leaf/measure, 
                                 filter(data, species == 1, individual == 1)), 1)
  lapply(list("species/individual" = var1,
              "individual/leaf" = var2, 
              "leaf/measure" = var3),
         function(var) 
           data.frame(level = names(var), variance = as.vector(var))) %>% 
    bind_rows(.id = "comparison") %>% 
    mutate(level = factor(level, levels = c("species", "individual", "leaf",
                                            "measure", "Within"))) %>% 
    mutate(comparison = factor(comparison, levels = c("species/individual", "individual/leaf",
                                                      "leaf/measure")))
}
lapply(list("full" = data, "sampling" = subdata), fun) %>% 
  bind_rows(.id = "data") %>% 
  ggplot(aes(x = data, y = variance, fill = level)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ comparison)
```
