```{r setupinds, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(V.PhyloMaker)
library(ggfortify)
library(ggtree)
library(sf)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Individuals

This chapter define sampled species and individuals.

**_Protium stevensonii_ is not in my Paracou DB !!** 

## Species

We selected 10 species (Tab. \@ref(tab:speciesTab)) common with the METRADICA project representative of the tree phylogeny in Paracou (Fig. \@ref(fig:speciesPhylo)).

```{r speciesTab}
species <- data.frame(taxon = c(
  "Casearia_javitensis", 
  "Chrysophyllum_prieurii", 
  "Conceveiba_guianensis",
  "Gustavia_hexapetala",
  "Jacaranda_copaia subsp. copaia",
  "Laetia_procera",
  "Protium_stevensonii",
  "Tachigali_melinonii",
  "Virola_michelii",
  "Virola_surinamensis"
)) %>% 
  separate(taxon, c("Genus", "Species"), remove = F) 
kable(select(species, Genus, Species), caption = "Studied species.")
```

```{r speciesPhylo, fig.cap="Selected species phylogeny."}
paracou <- read_csv2("data/Paracou_P16_2020.csv") %>% 
  dplyr::select(Family, Genus, Species) %>% 
  unique() %>% 
  left_join(species) %>% 
  mutate(species = paste(Genus, Species), genus = Genus, family = Family) %>% 
  dplyr::select(species, genus, family, taxon) %>% 
  mutate(taxon = as.character(taxon))
paracou <- mutate(paracou, taxon = ifelse(species == "Protium Indet.", "Protium", taxon)) # for the missing Protium
# tree <- phylo.maker(sp.list = paracou, tree = GBOTB.extended, nodes = nodes.info.1, scenarios = "S3")
# save(tree, file = "save/phylogeny.Rdata")
load("save/phylogeny.Rdata")
fortify(tree$scenario.3) %>% 
  mutate(species = gsub("_", " ", label)) %>% 
  left_join(paracou) %>% 
  ggtree(aes(col = taxon), layout="circular") + 
  geom_tiplab2(aes(alpha = !is.na(taxon), size = !is.na(taxon))) +
  theme_tree(legend.position='right', legend.text = element_text(face = "italic")) +
  scale_alpha_manual("taxon", values = c(0.2, 1)) +
  scale_size_manual("taxon", values = c(1, 2))
```

## Individuals from METRADICA

> Exclude Marion's individuals (Tab. \@ref(tab:indMD)) and collected or wrong individuals (Tab. \@ref(tab:indMB)) 

```{r indMD}
readxl::read_excel("data/202012_SelectionIndividus_V5.4.xlsx", sheet = "IndAMesurer") %>% 
  filter(TaxonActu %in% species$taxon) %>% 
  mutate(Statut = recode(Statut,
                         "AMesurerMD" = "Candidate",
                         "AMesurerMDPrioDK" = "Candidate",
                         "MesuréDK" = "Measured",
                         "MesuréDrought" = "Measured",
                         "MesuréMD" = "Measured",
                         "MesuréMD (MauvaiseSp)" = "Wrong",
                         "MesuréMD+Drought" = "Measured",
                         "MortMD" = "Wrong",
                         "MortParacou20" = "Wrong",
                         "MortParacou20PrioDK" = "Wrong",
                         "PasFaisable" = "Wrong",
                         )) %>% 
  group_by(TaxonActu, Statut) %>% 
  summarise(N = n()) %>% 
  reshape2::dcast(TaxonActu ~ Statut, value.var = "N") %>% 
  mutate(TaxonActu = gsub("_", " ", TaxonActu)) %>% 
  kable(caption = "Sumary of individuals (from 202012_SelectionIndividus_V5.4.xlsx)")
```

```{r indMB}
readxl::read_excel("data/FTH2021_SelectionIndividus_V5.5.xlsx", sheet = "IndAMesurer") %>% 
  filter(TaxonActu %in% species$taxon) %>% 
  group_by(TaxonActu, Plot) %>% 
  summarise(N = n()) %>% 
  reshape2::dcast(TaxonActu ~ Plot, value.var = "N") %>% 
  mutate_all(funs(ifelse(is.na(.), " ", .))) %>% 
  kable(caption = "Sumary of candidates (from FTH2021_SelectionIndividus_V5.5.xlsx)")
```

## Diameters

> minimizing DBH (Tab. \@ref(fig:dbhdist))

```{r dbhdist, fig.cap="DBH distribution for selected species in the plot 16 of Paracou in 2020."}
read_csv2("data/Paracou_P16_2020.csv") %>% 
  left_join(species) %>% 
  filter(!is.na(taxon)) %>% 
  mutate(DBH = CircCorr/pi) %>% 
  ggplot(aes(DBH)) +
  geom_histogram() +
  facet_wrap(~ paste(Genus, Species), scales = "free")
```

## Topography

> maximizing TWI variation (Tab. \@ref(fig:twidist))

```{r twidist, fig.cap="TWI distribution for selected species in the plot 16 of Paracou in 2020."}
paracou <- read_csv2("data/Paracou_P16_2020.csv") %>% 
  left_join(species) %>% 
  filter(!is.na(taxon)) %>% 
  mutate(DBH = CircCorr/pi) %>% 
  st_as_sf(coords = c("Xutm", "Yutm"),
           crs = '+proj=utm +zone=22 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0')
paracou$TWI <- raster::extract(raster::raster("data/TWI_1m.tif"), paracou)
ggplot(paracou, aes(TWI)) +
  geom_histogram() +
  facet_wrap(~ paste(Genus, Species), scales = "free")
```


## Candidates

> 15-20 individuals
